---
title: "NYPD Shooting Incident Data Report"
author: "T.M"
date: "2023-08-21"
output:
  pdf_document: default
  html_document: default
---

* According to NYC Open data, this data is a breakdown of every shooting incident that occurred in NYC from 2006 to the end of the previous calendar year.

* This paper will describe the steps to load data, perform a simple analysis, and build a simple model.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1.import libraries

* First, import the libraries to be used in this R program.

```{r import libraries}
library("tidyverse")
library("lubridate")
```


## 2.load data

* Next, download the data directly from the city of new york website and check the contents of the data.

```{r load_data}

# load the data
url <- "https://data.cityofnewyork.us/api/views/833y-fsy8/rows.csv?accessType=DOWNLOAD"
df <- read_csv(url)

# check the data
summary(df)
head(df,10)
colnames(df)
```

## 3. tidy data

* Then, drop unnecessary columns and check missing values.

```{r tidy_data}

# eliminate data which is not needed
df <- df %>% select(
  INCIDENT_KEY,
  OCCUR_DATE,
  OCCUR_TIME,
  BORO,
  LOC_OF_OCCUR_DESC,
  LOC_CLASSFCTN_DESC,
  LOCATION_DESC,
  STATISTICAL_MURDER_FLAG,
  PERP_AGE_GROUP,
  PERP_SEX,
  PERP_RACE,
  VIC_AGE_GROUP,
  VIC_SEX,
  VIC_RACE,
  Latitude,
  Longitude
)

# check missing values
missing_values <- df %>%
  summarise_all(function(x) sum(is.na(x)))
print(missing_values)
```
## 4. Visualization and Analysis

* Narrow down to important attributes and tabulate. Then, visualize them.

```{r Visualization and Analysis}
# aggregate data
boro_summary <- df %>%
  group_by(BORO) %>%
  summarise(count = n()) %>%
  filter(count > 1)

# plot bar chart
boro_plot <- ggplot(boro_summary, aes(x = BORO, y = count)) +
  geom_bar(stat = "identity") + 
  labs(title = "NUMBER OF INCIDENTS BY BORO",
       x = "BORO",
       y = "NUMBER OF INCIDENTS") +
  theme_minimal()

print(boro_plot)

# aggregate data
age_summary <- df %>%
  group_by(VIC_AGE_GROUP) %>%
  summarise(count = n()) %>%
  filter(count > 1)

# plot bar chart
age_plot <- ggplot(age_summary, aes(x = VIC_AGE_GROUP, y = count)) +
  geom_bar(stat = "identity") + 
  labs(title = "NUMBER OF INCIDENTS BY VIC_AGE_GROUP",
       x = "VIC_AGE_GROUP",
       y = "NUMBER OF INCIDENTS") +
  theme_minimal()

print(age_plot)


# aggregate data
sex_summary <- df %>%
  group_by(VIC_SEX) %>%
  summarise(count = n()) %>%
  filter(count > 1)

# plot bar chart
sex_plot <- ggplot(sex_summary, aes(x = VIC_SEX, y = count)) +
  geom_bar(stat = "identity") + 
  labs(title = "NUMBER OF INCIDENTS BY VIC_SEX",
       x = "VIC_SEX",
       y = "NUMBER OF INCIDENTS") +
  theme_minimal()

print(sex_plot)
```
The visualization revealed the following;

  * Incidents are more than 20% higher in Brooklyn than in the Bronx
  * The volume zone for age is 18 to 44. It drops off dramatically for those over 45.
  * The majority of incidents by gender are male.

## 5. Modeling

* Model the impact on STATISTICAL_MURDER_FLAG using key variables.

```{r modeling}
df_filtered <- df %>%
  filter(!is.na(STATISTICAL_MURDER_FLAG),
         !is.na(BORO),
         !is.na(VIC_AGE_GROUP),
         !is.na(VIC_SEX)) %>%
  select(STATISTICAL_MURDER_FLAG, BORO, VIC_AGE_GROUP, VIC_SEX)

dummy_data <- model.matrix(~ BORO + VIC_AGE_GROUP + VIC_SEX - 1, data = df_filtered)

df_combined <- as.data.frame(dummy_data) %>%
  bind_cols(df_filtered["STATISTICAL_MURDER_FLAG"])

model <- lm(STATISTICAL_MURDER_FLAG ~ ., data = df_combined)

summary(model)
```

* Variables that can be utilized when considering P-values are BOROMANHATTAN and VIC_AGE_GROUP.
* BORO = Manhattan has a negative effect on STATISTICAL_MURDER_FLAG, while VIC_AGE_GROUP seems to have a positive effect as one gets older.


# Comment
At first, I thought that the younger one was, the higher the risk, and that a higher STATISTICAL_MURDER_FLAG indicated increased danger. I also believed that there were more incidents in the Bronx than in Brooklyn. However, I realized that this was on a bias.
